Local-global principle for $\K_2$ of Chevalley groups

### 2015-09-30 paper.tex
Written down Lavrenov's computation

\begin{dfn}
 The relative Steinberg group $\St^*(n,R, I)$ is the group defined by the following two
 families generators and four families of relations.
 \begin{itemize}
  \item Generators:
  \begin{enumerate}
  \item $X^1(u, v)$, where $u \in \E(n,R) \cdot e_1$, $v\ \in I^n$ such that $v^t \cdot u = 0$;
  \item $X^2(u, v)$, where $u \in I^n$, $v \in \E(n,R) \cdot e_1$ such that $v^t \cdot u = 0$.
 \end{enumerate}
  Notice that the canonical image in $\E(n, R)$ of both $X^1(u, v)$ and $X^2(u, v)$ equals $T(u, v) = e_n + u \cdot v^T$.
  \item Relations:
  \begin{enumerate}
  \item $X^1(u, v) \cdot X^1(u, w) = X^1(u, v+w)$, $u \in \E(n,R) \cdot e_1$, $v, w \in I^n$;
  \item $X^2(u, v) \cdot X^2(w, v) = X^2(u+w, v)$, $u, w \in I^n$, $v \in \E(n,R) \cdot e_1$;
  \item ${}^{X^\sigma(u^2, v^2)} \! X^\tau(u^1, v^1) = X^\tau(T(u^2, v^2) \cdot u^1, T(v^2, u^2)^{-1} \cdot v^1)$, $\sigma, \tau = 1..2$;
  \item $X^1(u, vb) = X^2(ub, v)$ for $(u,v) \in E(n,R) \cdot (e_1, e_2)$, $b\in I$.
 \end{enumerate}
 \end{itemize}
\end{dfn}
It is easy to check the correctness of the above definition.

@@

The next step of the proof is to is construct certain elements in $\St(n, R)$ which are similar to Tulenbaev's elements $X_{u,v}(a)$ see~\cite[\S~1]{T}.

Assume that $n \geq 4$. For the rest of this section $a$ denotes a nonnilpotent element of $R$.
\begin{lemma}
 Let $u, v \in R^n$ be columns such that $v^t \cdot u = 0$ and there exists $w = w(u) \in R^n$ satisfying $w^t \cdot u = a^N$ for some natural number $N$.
 In this context one can define certain elements $Z^{1}(u,v), Z^{2}(v, u)$ of $\St(n, R)$. These elements satisfy the following relations:
 \begin{itemize}
  \item $Z^{1}(u, v_1 + v_2) = Z^{1}(u, v_1) \cdot Z^{1}(u, v_2)$;
  \item $Z^{2}(v_1 + v_2, u) = Z^{2}(v_1, u) \cdot Z^{2}(v_2, u)$;
  \item ${}^{g}\! Z^{\tau}(u, v) = Z^{\tau}(g \cdot u, g^* \cdot v)$, $\tau = 1..2$, $g \in \St(n, R)$ where $g^* = {g^t}^{-1}$ is the contragradient matrix.
 \end{itemize}
\end{lemma}
\begin{proof} Both the construction of $Z^i$ and the relations for $Z^i$ are, in essence, just~\cite[Lemma~1.3]{T} applied twice. \end{proof}

%TODO: Why we can assume u_1^t \cdot u_2 = 0$?
Now let $g \in \E(n, R_a)$ be such that $g(e_1, e_2) = (\overline{u_1}, \overline{u_2})$ for some $u_1, u_2\in R^n$, $u_1^t \cdot u_2 = 0$.
It is easy to find $w_1, w_2 \in R^n$ such that $w_1^t \cdot u_1 = a^{N_1}$, $w_2^t \cdot u_2 = a^{N_2}$.

$Z^1(u, v) = Z^2(v, u)$.

for each $(u,v) \in \E_n(R) (e_1, e_2)$ we can find $y$ such that $(u,v,y) = g (e_1, e_2, e_3)$.

$w^tu = a^N$, $z^Tv = a^N$, $x^T y = a^N$ + 6 additional orthogonality relations

Computing $g = [Z^{1, N}(u xba^{2N}), Z^{2, N}(ua^{2N},v)]$ 
$$ g = Z^{2, N}(ya^{2N} + uba^{5N},v) {Z^{2, N}(ya^{2N}, v)}^{-1} = Z^{2, N}(uba^{5N},v) = Z^2(uc, v)$$
$$ g = ... = Z^1(u, vc)$$

>>>>

\begin{dfn}
 The relative Steinberg group $\St^*(n,R, I)$ is the group defined by the following two
 families generators and four families of relations.
 \begin{itemize}
  \item Generators:
  \begin{enumerate}
  \item $X^1(u, v)$, where $u \in \E(n,R) \cdot e_1$, $v\ \in I^n$ such that $v^t \cdot u = 0$;
  \item $X^2(u, v)$, where $u \in I^n$, $v \in \E(n,R) \cdot e_1$ such that $v^t \cdot u = 0$.
 \end{enumerate}
  Notice that $\phi$ maps both $X^1(u, v)$ and $X^2(u, v)$ to $T(u, v) = e_n + u \cdot v^T$.
  \item Relations:
  \begin{enumerate}
  \item $X^1(u, v) \cdot X^1(u, w) = X^1(u, v+w)$, $u \in \E(n,R) \cdot e_1$, $v, w \in I^n$;
  \item $X^2(u, v) \cdot X^2(w, v) = X^2(u+w, v)$, $u, w \in I^n$, $v \in \E(n,R) \cdot e_1$;
  \item ${}^{X^\sigma(u^2, v^2)} \! X^\tau(u^1, v^1) = X^\tau(T(u^2, v^2) \cdot u^1, T(v^2, u^2)^{-1} \cdot v^1)$, $\sigma, \tau = 1..2$;
  \item $X^1(g \cdot e_1, g^* \cdot be_2) = X^2(g \cdot be_1, g^* \cdot e_2)$ where $b\in I$ and $g^* = {g^t}^{-1}$ denotes the contragradient matrix.
 \end{enumerate}
 \end{itemize}
\end{dfn}
It is easy to check the correctness of the above definition.

@@

The next step of the proof is to is construct certain elements in $\St(n, R)$ similar to Tulenbaev's elements $X_{u,v}(a)$ see~\cite[\S~1]{T}.

Let $v\in R^n$ be a column.
A column $w\in R^n$ is called \emph{$v$-decomposable} if it can be presented as a finite sum $w = \sum\limits_{i=1}^n w^i$ such that each $w^i$ has at least two zero entries. 
Denote by $D(v)$ the submodule of $R^n$ consisting of all $v$-decomposable columns.
For a column $v\in R^n$ denote by $I(v)$ the ideal of $R$ spanned by coordinates $v_1,\ldots, v_n$.

Let $u,v,w\in R^n$ be columns such that $w^tv=0$.
It is easy to check (cf.~\cite[Lemma~3.2]{Ka}) that 
$$(uv)\cdot w = \sum_{i<j}w_{ij},\ \text{where}\ w_{ij} = (w_iu_j - w_ju_i)(v_je_i - v_ie_j)\in{}\!A^n.$$
The above decomposition is called the \emph{canonical} decomposition of $(uv)\cdot w$.
In particular, the above argument shows that row $a\cdot w$ is always $v$-decomposable for $a\in I(v)$, i.e. $I(v) \cdot O(v) \subseteq D(v)$.
It is also obvious that $O(v)\subseteq O(bv)$, $D(v)\subseteq D(bv)$ for $b \in R$.
%\end{enumerate}\end{lemma}

Denote by $B^1$ the subset of $R^n \times R^n \times R$ consisting of triples $(u, v, a)$ such that $v^t \cdot u = 0$, $v \in D(u)$, $a \in I(u)$.
Denote by $B^2$ the set consisting of triples $(v, u, a)$ such that $(u, v, a) \in B^1$.

\begin{lemma} \label{lem:Zproperties}
Assume that $n \geq 4$.
One can define two families of elements $Z^\tau(u, v, a) \in \St(n, R)$, $\tau=1,2$ parametrized by $(u, v, a) \in B^\tau$ satisfying the following properties
 \begin{lemlist}
  \item \label{rel:proj} $\phi(Z^\tau(u, v, a)) = e + uav^t \in \E(n, R)$, $(u,v,a) \in B^\tau$;
  \item \label{rel:add1} $Z^{1}(u, v_1 + v_2, a) = Z^{1}(u, v_1, a) \cdot Z^{1}(u, v_2, a)$;
  \item \label{rel:add2} $Z^{2}(v_1 + v_2, u, a) = Z^{2}(v_1, u, a) \cdot Z^{2}(v_2, u, a)$;
  \item \label{rel:conj} ${}^{g}\! Z^{\tau}(u, v, a) = Z^{\tau}(\phi(g) \cdot u, \phi(g)^* \cdot v, a)$, $\tau = 1,2$, $g \in \St(n, R)$.
 \end{lemlist}
\end{lemma}
\begin{proof}
One constructs the elements $Z^1(u,v,a)$ in exactly the same way as Tulenbaev constructs his elements $X_{u,v}(a)$ (see definitions preceding~\cite[Lemma~1.2]{T}).
Next, the elements $Z^2(u,v,a)$ are constructed by ``transposing'' Tulenbaev's construction.

All the properties (with the exception of the last one in the case $n=4$) can be proved in exactly the same fashion as in~\cite[Lemma~1.3]{T}.
To check the last property in the case $n=4$ one has to use the argument of W. van~der~Kallen (cf.~\cite{Ka}).
%TODO: Write more details
\end{proof}

For the rest of this section $a$ denotes a nonnilpotent element of $R$.
\begin{lemma}
Let $u,v,w,z \in R^n$ be columns such that 
\begin{equation*} u^t \cdot v = 0,\ w^t \cdot v = c \in R,\ u^t \cdot z = 0,\ w^t \cdot z = 0. \end{equation*}
Assume, moreover, that $(u, v, a) \in B^1$, $(w, z, a) \in B^2$, then one has $$Z^1(u, ac \cdot z, a) = Z^2(ac \cdot u, z, a).$$
\end{lemma}
\begin{proof}
Direct computation using \cref{lem:Zproperties} shows that
 \begin{multline*}
 Z^2(ac \cdot u, z, a) = Z^2((e+uav^t) w, (e-vau^t)z,a) \cdot Z^2(-w,z,a) = \\
  {}^{Z^1(u,v,a)} \! Z^2(w,z,a) \cdot Z^2(-w,z,a) = [Z^1(u, v, a), Z^2(w, z, a)] = Z^1(u,v,a) \cdot {}^{Z^2(w,z,a)}\! Z^1(u,-v,a) = \\
    = Z^1(u,v,a) \cdot Z^1((e+waz^t)u,-(e-zaw^t)v,a) = Z^1(u,ac \cdot z,a). \qedhere
 \end{multline*}  
\end{proof}
